Bandit Multiclass Classification

### 2018-10-11 ptf_margin.tex
added elaboration on norm bound

%\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
% For citations
\usepackage{natbib}
\usepackage{amsthm}
\usepackage{makecell}
% For algorithms
%\usepackage[algcompatible]{algpseudocode}
%\usepackage{algorithm}
%\usepackage{algorithmicx}
%\usepackage[noend]{algcompatible}
\usepackage[vlined,linesnumbered,ruled]{algorithm2e}
\usepackage{bbm}
\usepackage{amsfonts}
\usepackage{amsmath}           
\usepackage{mathtools}    
\usepackage{amsopn}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{soul}
%%%%%%%%%%%%%%%%%% toc 
\usepackage{scrwfile}
\TOCclone[\contentsname~(\appendixname)]{toc}{atoc}
\newcommand\StartAppendixEntries{}
\AfterTOCHead[toc]{%
  \renewcommand\StartAppendixEntries{\value{tocdepth}=-10000\relax}%
}
\AfterTOCHead[atoc]{%
  \edef\maintocdepth{\the\value{tocdepth}}%
  \value{tocdepth}=-10000\relax%
  \renewcommand\StartAppendixEntries{\value{tocdepth}=\maintocdepth\relax}%
}
\newcommand*\appendixwithtoc{%
  \cleardoublepage
  \appendix
  \addtocontents{toc}{\protect\StartAppendixEntries}
  \listofatoc
}
\usepackage{blindtext}
\usepackage{color}
%%%%%%%%%%%%%%%%%% toc

@@

\DeclareMathOperator*{\arginf}{arginf}
\DeclareMathOperator*{\argsup}{argsup}
\DeclareMathOperator*{\range}{range}
\DeclareMathOperator*{\mydet}{det_{+}}
\DeclarePairedDelimiter\abs{\lvert}{\rvert}
\DeclarePairedDelimiter\bigabs{\big\lvert}{\big\rvert}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\DeclarePairedDelimiter\bigceil{\big\lceil}{\big\rceil}
\DeclarePairedDelimiter\bigfloor{\big\lfloor}{\big\rfloor}

@@

\newcommand{\theset}[2]{ \left\{ {#1} \,:\, {#2} \right\} }
\newcommand{\inner}[1]{ \left\langle {#1} \right\rangle }
\newcommand{\Ind}[1]{ \field{I}_{\{{#1}\}} }
\newcommand{\eye}[1]{ \boldsymbol{I}_{#1} }
\newcommand{\norm}[1]{\left\|{#1}\right\|}
%\newcommand{\trace}[1]{\text{tr}\left({#1}\right)}
\newcommand{\trace}[1]{\textsc{tr}({#1})}
\newcommand{\diag}[1]{\mathrm{diag}\!\left\{{#1}\right\}}

@@

\newtheorem{assumption}{Assumption}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\DeclareMathOperator{\val}{val}
\DeclareMathOperator{\spa}{sp}
\DeclareMathOperator{\solve}{solve}
\allowdisplaybreaks

@@

\section{Problem Statement}
A multiclass linear classifier is defined by a set of $K$ vectors $w^{(1)}, \ldots, w^{(K)}\in \mathbb{R}^d$ (or a compact matrix form $W = [w^{(1)}\ w^{(2)}\  \cdots\  w^{(K)}]^\top \in \mathbb{R}^{K\times d}$), where $K$ is the number of classes ($K\geq 3$) and $d$ is the dimension of the feature vector. A set of samples $\{(x_t, y_t)\}_{t=1}^T$ where $x_t\in \mathbb{R}^d$ and $y_t\in[K]$ is called \textit{linearly separable} if there exists a $w_*^{(1)}, \ldots, w_*^{(K)}$ (or simply $W_*$) such that $ y_t= \argmax_{y\in[K]} (W_{*}x_t)_y $ for all $t\in[T]$.

A set of samples $\{(x_t, y_t)\}_{t=1}^T$ is called \textit{linearly separable with a margin $\gamma$} if there exists $W_*$ with $\|w^{(i)}_*\|_2\leq 1$ for all $i$ and $(W_{*}x_t)_{y_t} \geq \max_{y\neq y_t} (W_*x_t)_y + \gamma$ for all $t$. We define the set $\calW = \{W\in \mathbb{R}^{K\times d}: \|\e_i^\top W^*\|_2\leq 1\}$. Therefore, $W_*\in \calW$.

Consider the following \textit{online multiclass classification} problem: at time $t$, the environment first reveals $x_t\in \mathbb{R}^d$ (and we assume $\|x_t\|_2\leq 1$), then the learner predicts some label $\tilde{y}_t\in [K]$; finally the environment reveals the true label $y_t\in[K]$. It is known that if the samples are linearly separable, then the learner will only make constant mistakes. It is achievable with the following simple perceptron algorithm:

\begin{algorithm}[H]
\caption{Perceptron}
$W_1=\mathbf{0}$\\
\For{$t=1$ to $T$}{
   Predict $\tilde{y}_t=\argmax_y (W_tx_t)_y$ \\
   Update $W_{t+1}\leftarrow W_t + (\e_{y_t}-\e_{\tilde{y}_t})x_t^\top$ \\
   (same as updating $w^{(y_t)}_{t+1} \leftarrow w^{(y_t)}_t + x_t$ and $w^{(\hat{y}_t)}_{t+1} \leftarrow w^{(\hat{y}_t)}_t - x_t$ when $\hat{y}_t\neq y_t$). 
}
\end{algorithm}
The analysis is simple. On one hand we have (note $\langle A,B \rangle:=\text{Tr}(A^\top B)$):
\begin{align*}
    \langle W_t, W_* \rangle &= \langle W_{t-1}, W_* \rangle + \langle (\e_{y_t}-\e_{\tilde{y}_t})x_t^\top, W_* \rangle \\
    &= \langle W_{t-1}, W_* \rangle + (\e_{y_t}^\top W_* x_t-\e_{\tilde{y}_t}^\top W_* x_t) \\
    &\geq \langle W_{t-1}, W_* \rangle + \gamma\one[\tilde{y}_t\neq y_t]. 
\end{align*}
By induction, we get $\langle W_{T+1}, W_* \rangle \geq \gamma\sum_{t=1}^T \one[\tilde{y}_t\neq y_t]$.

On the other hand, 
\begin{align*}
    \langle W_{T+1}, W_* \rangle \leq \|W_{T+1}\|_F\|W_*\|_F \leq \sqrt{2K\sum_{t=1}^T \one[\tilde{y}_t \neq y_t] \|x_t\|_2^2}\times 1 \leq \sqrt{2K\sum_{t=1}^T \one[\tilde{y}_t \neq y_t] } 
\end{align*}
Combining the above two inequality, we get $ \sum_{t=1}^T \one[\tilde{y}_t \neq y_t] \leq \frac{4K}{\gamma^2}$.

The \textit{bandit multiclass classification} problem proceeds in a very similar way. The only difference is that the learner only knows whether she predicts \textbf{correctly or not}, rather than gets the true label (i.e., in each round, the feedback is only $\one[\tilde{y}_t=y_t]$, rather than $y_t$). Surprisingly, with this extremely limited feedback, the learner can still make only constant mistakes. It can be achieved with the following variant of \textit{halving algorithm}:

\begin{algorithm}[H]
\caption{Halving}
Discretize the space of $W$ (i.e., the $\calW$ as defined in the beginning) with balls of radius $\frac{1}{2\gamma}$. Let the set of discretization points be $\calS$. Then this is saying that for all $W\in \calW$, there is always a $W'\in \calS$ such that $\|W-W'\|_F\leq \frac{1}{2\gamma}$.\\
Let $\calS_1=\calS$ \\
\For{$t=1$ to $T$}{
    Let $\calS_t(y) = \{W\in \calS_t: (Wx_t)_y \geq (Wx_t)_i \ \forall i\in[K]\}$ (i.e., the set of $W$'s in $\calS_t$ that predict $y$ as the label of $x_t$) \\
    Let $\tilde{y}_t = \argmax_y |\calS_t(y)|$ (i.e., pick the majority vote)\\
    \If{$\tilde{y}_t\neq y_t$}{
        $\calS_{t+1}=\calS_t \backslash \calS_t(\tilde{y}_t)$ (i.e., eliminate the $W$'s that are inconsistent with the outcomes)
    }
    \Else{
        $\calS_{t+1}=\calS_t$
    }
}
\end{algorithm}
\textbf{Analysis}.  The majority in $\calS_t$ always has cardinality no less than $\frac{1}{K}|\calS_t|$. So every time the learner predicts incorrectly, the size of $|\calS_t|$ shrinks by an order of $(1-\frac{1}{K})$. By our margin assumption, there is a $W'\in \calS$ which incurs no error in all samples. Therefore, $|S_t|\geq 1$ always holds. This means the number of errors is bounded by the order of $\frac{\log(|\calS|)}{-\log(1-\frac{1}{K})}\leq \mathcal{O}\left( K^2d\log \frac{1}{\gamma} \right)$.

The main issue of this halving algorithm is that it is \textbf{inefficient} in the sense that $|\calS|$ is in the order of $\frac{1}{\gamma^{Kd}}$.

Hence, we want to answer the following question: \ul{for bandit multiclass classification, can we have an efficient algorithm (time complexity polynomial in $K,d,\frac{1}{\gamma}, T$) that guarantees constant mistake bound (polynomial in $K,d,\frac{1}{\gamma}$) in the $\gamma$-separable case?} (An unsolved open problem in \cite{kakade2008efficient})

\section{Naive Approaches}
\begin{itemize}
    \item \textbf{Exhausting the feature space}: discretize the feature space into $ \left(\frac{1}{\gamma}\right)^{\mathcal{O}(d)} $ blocks. The margin assumption guarantees that two points in a block would have the same label (so we can give every block a label). Each time a new point $x_t$ comes, see whether the learner already knows the correct label for that block. If yes, predict that label; if not, randomly choose a label that has not been chosen for that block. \\
    $\Rightarrow $ error bound: $K\left(\frac{1}{\gamma}\right)^{\mathcal{O}(d)} $. 
    \item \textbf{Exhausting the hypothesis space}: Discretize the hypothesis space into $\left(\frac{1}{\gamma}\right)^{Kd}$ discretization points (like in the halving algorithm). For each of them, make predictions based on it until it makes an mistake. \\
    $\Rightarrow $ error bound: $\left(\frac{1}{\gamma}\right)^{Kd}$. 
\end{itemize}

\section{Difficulties}
\label{section: difficulty}
Efficient classification algorithms often optimize a convex surrogate loss rather than a 0-1 loss. Examples include: 
\begin{itemize}
    \item Hinge loss: $\ell_t(W) = [1-(Wx_t)_{y_t}+\max_{y\neq y_t}(Wx_t)_{y}]_+$. 
    \item Logistic loss: $\ell_t(W)=-\ln\frac{\exp((Wx_t)_{y_t})}{\sum_{y}\exp((Wx_t)_y)}$
    \item A family of loss that interpolates hinge loss and squared hinge loss (see \cite{beygelzimer2017efficient}'s Eq.(3))
\end{itemize}
In full-information setting, one can directly use online convex optimization techniques to deal with the above loss functions (learning over the space of $W$).

Many works for bandit classification reuse this kind of convex optimization schemes, together with explicit exploration and inverse propensity weighting \cite{kakade2008efficient, hazan2011newtron, beygelzimer2017efficient, foster2018logistic}. However, because it is hard to estimate these losses when the true label $y_t$ is not known, some of these algorithms \cite{beygelzimer2017efficient, foster2018logistic} simply do not update when the learner makes a mistake ($\tilde{y}_t \neq y_t$). We give this kind of algorithm an error lower bound $\Omega\left(\left(\frac{1}{\gamma}\right)^{(d-1)/2}\right)$ in Section~\ref{section: passive lower bound}.

It indeed looks hard to design a convex loss for $W$'s when the learner makes a mistake: when $\tilde{y}_t
\neq y_t$, the set of $W$'s that we want to penalize (i.e., to assign larger loss) is $\{W: (Wx_t)_{\tilde{y}_t} \geq (Wx_t)_{y}, \forall y \}$, which is a convex cone in the space of $W$. It is impossible for a convex function to be large only in a convex subset (the case $K=2$, on the other hand, does not have this issue). Can we argue that if the learner is restricted to use convex losses in the space of $W$, she will have to suffer exponential (in $K$) errors?

\section{One-versus-all Perceptron}
\label{sec:ova}
Our assumption of linearly separable with a margin is 
\begin{align}
 (W_*x_t)_{y_t} \geq (W_*x_t)_{y} + \gamma, \ \forall y\neq y_t. \label{eqn:weaker}
\end{align}
As discussed in Section~\ref{section: difficulty}, it seems hard to make update when $\tilde{y}_t\neq y_t$ and achieve a constant and polynomial error bound.

For a moment, in this section we make the following stronger margin assumption (which we call \textit{one-versus-all separable} with a margin): 
\begin{align}
\begin{cases}
    (W_*x_t)_{y_t} \geq \gamma/2 \\
    (W_*x_t)_y \leq -\gamma/2, \forall y\neq y_t.
\end{cases}
\label{eqn:stronger}
\end{align}
Clearly, one-versus-all separability implies linearly separability, but not the other way around. With one-versus-all separability assumption, we can view the problem as $K$ parallel binary classification problem. The following algorithm achieves constant error bound: 
\begin{algorithm}[H]
\caption{One-versus-all Perceptron}
\label{alg:ova}
Initialize $w_t^{(1)}=\ldots=w_t^{(K)}=\mathbf{0}\in \mathbb{R}^d$\\
\For{$t=1$ to $T$}{
   \If{$\exists y$ such that $w_t^{(y)\top} x_t \geq 0$}{
       Assign $\tilde{y}_t$ to any $y$ with $w_t^{(y)\top}  x_t \geq 0$ \\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t\neq y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} - x_t$ } \label{line:update1}
   }
   \Else{
       Pick $\tilde{y}_t$ randomly from $
       \text{unif}[K]$ \label{line:explore}\\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t = y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} + x_t$ } \label{line:update2}
   }
}
\end{algorithm}
\textbf{Analysis}. Note that when the algorithm enters Line~\ref{line:update1} or Line~\ref{line:update2}, the binary classifier $w_t^{(\tilde{y}_t)}$ is making an error. Let the number of times the algorithm enters Line~\ref{line:update1} and Line~\ref{line:update2} be $M$ and $N$ respectively. By the error bound of binary perceptron $\mathcal{O}\left(\frac{1}{\gamma^2}\right)$, we have that $M+N=\mathcal{O}\left(\frac{K}{\gamma^2}\right)$. Then note that the number of times the algorithm makes a mistake (i.e. $\tilde{y}_t\neq y_t$) is upper bounded by $M$ plus how many times the algorithm explores in Line~\ref{line:explore}; and when the algorithm explores, with probability $\frac{1}{K}$ it enters Line~\ref{line:update2}. Therefore, the number of mistakes is bounded (in expectation) by $\E[M+KN]=\mathcal{O}\left(\frac{K^2}{\gamma^2}\right)$.

\section{One-versus-all Kernel Perceptron}
%With the polynomial kernel defined in \cite{klivans2004learning}, we can transform the weaker assumption \eqref{eqn:weaker} in the original feature space to the stronger one \eqref{eqn:stronger} in a transformed feature space. Indeed, in the original feature space, class $i$ corresponds to an intersection of $K-1$ halfspaces: $ \{x: (w_i^*-w_j^*)^\top x \geq 0, \forall j\neq i\} $, which is the subject of \cite{klivans2004learning} (here, $w_i^{*\top}$ equals to $\e_i^\top W^*$ defined above). Directly using their kernel construction, we can have margin $ \left(\frac{\gamma}{d}\right)^{\Omega(K\log K\log (1/\gamma))} $ in the transformed feature space. Running kernelized version of Algorithm~\ref{alg:ova}, we can get $ K^2\left(\frac{d}{\gamma}\right)^{\mathcal{O}(K\log K\log (1/\gamma))} $ error bound directly. 
The idea here is to do a feature transformation such that if the original assumption \eqref{eqn:weaker} is satisfied, then in the transformed feature space, \eqref{eqn:stronger} will be satisfied, then we can run the one-versus-all perceptron given in \ref{sec:ova}. Under the original assumption, each class corresponds to an intersection of $K-1$ halfspaces in the feature spaces (i.e., class $i$ corresponds to $\{x\in\mathbb{R}^d: w_*^{(i)\top}x \geq w_*^{(j)\top}x, \ \forall j\neq i\}$). We want this region to become a halfspace in the transformed space. Fortunately, this is possible when there is a margin, and is exactly what is done in \cite{klivans2004learning}.

\subsection{Refinement for \cite{klivans2004learning}'s Fact 1}
\begin{lemma}
   For $i=1,\ldots, \ell$ let $q^{(i)}(x)=\sum_S c^{(i)}_{S} x_S $ be a polynomial over $x_1, \ldots, x_d$ with $\|q^{(i)}\|^2\leq M_i$. Then (1) if $q^{(1)}\ldots q^{(\ell)}$ has degree at most $\deg$, we have $\|q^{(1)}\ldots q^{(\ell)}\|^2 \leq \ell^{\deg} \prod_i M_i$, and (2) we have $\|q^{(1)}+\ldots+q^{(\ell)}\|^2\leq \ell(M_1+\ldots+M_\ell)$. 
\end{lemma}
\begin{proof}
For the first bound, we bound the ratio between the following two values:  $ \|q^{(1)}\ldots q^{(\ell)}\|^2 $ and $\|q^{(1)}\|^2\ldots \|q^{(\ell)}\|^2$. 
\begin{align}
    n_1&=\prod_{i=1}^\ell \|q^{(i)}\|^2 
    = \prod_{i=1}^\ell \left(\sum_{S_i} \left(c^{(i)}_{S_i}\right)^2 \right)
    = \sum_{(S_1,\ldots,S_\ell)}  \left(\prod_{i=1}^\ell \left(c^{(i)}_{S_i}\right)^2 \right)
    = \sum_{(S_1,\ldots,S_\ell)}  \left(\prod_{i=1}^\ell c^{(i)}_{S_i}\right)^2 \nonumber\\
    &= \sum_S \sum_{\substack{(S_1,\ldots,S_\ell): \\ S_1\times\ldots S_\ell=S}}  \left(\prod_{i=1}^\ell c^{(i)}_{S_i}\right)^2
\end{align}
\begin{align}
    n_2=\|q^{(1)}\ldots q^{(\ell)}\|^2 = \sum_{S} c_S^2 
    = \sum_S \left( \sum_{\substack{(S_1,\ldots, S_\ell): \\ S_1\times\ldots\times S_\ell=S }} \left(c_{S_1}^{(1)}\ldots c_{S_\ell}^{(\ell)} \right) \right)^2 
    = \sum_S \left( \sum_{\substack{(S_1,\ldots, S_\ell): \\ S_1\times\ldots\times S_\ell=S }} \prod_{i=1}^\ell c_{S_i}^{(i)} \right)^2 
\end{align}
Let $M$ be an upper bound of the number of terms involved in the summation $ \sum_{\substack{(S_1,\ldots, S_\ell): \\ S_1\times\ldots\times S_\ell=S }} $, then by Cauchy-Schwarz's inequality we have $ n_2\leq Mn_1 $.

$M$ counts the number of different $(S_1,\ldots, S_\ell)$'s with $S_1\times \ldots \times S_\ell=S$. Since $S$ has degree at most $\deg$, we can bound $M$ by $ \ell^{\deg}$.

The second bound can be obtained by applying Cauchy-Schwarz once.

\end{proof}

\subsection{Refinement for \cite{klivans2004learning}'s Theorem 10}
\begin{lemma}
Under the same condition as in \cite{klivans2004learning}'s Theorem 10, the PTF has margin on $X$ is at least $(1/t)^{O(r\log r + r\log\log t)}$. 
\end{lemma}
\begin{proof}
We simply follow the construction in the original paper. $\|1-w^i\cdot x\|^2\leq 4$ can imply $\|(1-w^i\cdot x)^j\|^2\leq j^{O(j)}4^j \leq j^{O(j)}$. Then $\|a_j(1-w^i\cdot x)^j\|^2\leq 2^{2r}r^{O(r)}=r^{O(r)}$ follows. Thus $\|T_r(1-w^i\cdot x)\|^2=\|\sum_{j=0}^r a_j (1-w^i\cdot x)^j\|^2 \leq (r+1)^2 r^{O(r)} = r^{O(r)}$. Then, $\left(P(w^i\cdot x)\right)^{\log 2t}\leq (\log t)^{O(r\log t)}\times r^{O(r\log t)}\leq (r\log t)^{O(r\log t)}$. Finally, $\|p\|^2\leq (t+1)^2 (r\log t)^{O(r\log t)} \leq (t+1)^2 \left(  t^{O(r\log r + r\log\log t)}\right) = t^{O(r\log r + r\log\log t)}$.  
\end{proof}

>>>>

%\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
% For citations
\usepackage{natbib}
\usepackage{amsthm}
\usepackage{makecell}
% For algorithms
%\usepackage[algcompatible]{algpseudocode}
%\usepackage{algorithm}
%\usepackage{algorithmicx}
%\usepackage[noend]{algcompatible}
\usepackage[vlined,linesnumbered,ruled]{algorithm2e}
\usepackage{bbm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amsopn}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{soul}
\usepackage{commath}
%%%%%%%%%%%%%%%%%% toc
\usepackage{scrwfile}
\TOCclone[\contentsname~(\appendixname)]{toc}{atoc}
\newcommand\StartAppendixEntries{}
\AfterTOCHead[toc]{%
  \renewcommand\StartAppendixEntries{\value{tocdepth}=-10000\relax}%
}
\AfterTOCHead[atoc]{%
  \edef\maintocdepth{\the\value{tocdepth}}%
  \value{tocdepth}=-10000\relax%
  \renewcommand\StartAppendixEntries{\value{tocdepth}=\maintocdepth\relax}%
}
\newcommand*\appendixwithtoc{%
  \cleardoublepage
  \appendix
  \addtocontents{toc}{\protect\StartAppendixEntries}
  \listofatoc
}
\usepackage{blindtext}
\usepackage{color}
%%%%%%%%%%%%%%%%%% toc

@@

\DeclareMathOperator*{\arginf}{arginf}
\DeclareMathOperator*{\argsup}{argsup}
\DeclareMathOperator*{\range}{range}
\DeclareMathOperator*{\mydet}{det_{+}}
%\DeclarePairedDelimiter\abs{\lvert}{\rvert}
\DeclarePairedDelimiter\bigabs{\big\lvert}{\big\rvert}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\DeclarePairedDelimiter\bigceil{\big\lceil}{\big\rceil}
\DeclarePairedDelimiter\bigfloor{\big\lfloor}{\big\rfloor}

@@

\newcommand{\theset}[2]{ \left\{ {#1} \,:\, {#2} \right\} }
\newcommand{\inner}[1]{ \left\langle {#1} \right\rangle }
\newcommand{\Ind}[1]{ \field{I}_{\{{#1}\}} }
\newcommand{\eye}[1]{ \boldsymbol{I}_{#1} }
%\newcommand{\norm}[1]{\left\|{#1}\right\|}
%\newcommand{\trace}[1]{\text{tr}\left({#1}\right)}
\newcommand{\trace}[1]{\textsc{tr}({#1})}
\newcommand{\diag}[1]{\mathrm{diag}\!\left\{{#1}\right\}}

@@

\newtheorem{assumption}{Assumption}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{fact}[theorem]{Fact}
\DeclareMathOperator{\val}{val}
\DeclareMathOperator{\spa}{sp}
\DeclareMathOperator{\solve}{solve}
\allowdisplaybreaks

@@

\section{Learning intersection of halfspaces with a margin}
Suppose we are given examples $S = \cbr{(x, y)}$, where $x$ is in $\cbr{x \in \fR^d: \| x \| \leq 1}$, and $y$ is in $\cbr{-1,+1}$.
\begin{definition}
$S$ is said to be separable by an intersection of halfspaces with margin parameter $\gamma > 0$, if there exists $t$ linear classifiers $v_1, \ldots, v_t$ (all of which are in $\cbr{v \in \fR^d: \| v \| \leq 1}$),
such that for every $(x,y)$ in $S$:
\begin{enumerate}
  \item If $y = +1$, then $v_i \cdot x \geq \gamma$ holds simultaneously for all $i$ in $[t]$.
  \item If $y = -1$, then there exists an $i$ in $[t]$, $v_i \cdot x \leq -\gamma$.
\end{enumerate}
\label{def:int-margin}
\end{definition}

Given a set $S$ separable by an intersection of halfspaces with a margin $\gamma$, our goal is to find a feature map $\phi: \fR^d \to \fR^{d'}$, such that the dataset $S' = \cbr{(\phi(x), y): (x,y) \in S}$ is linearly separable by a margin $\gamma' > 0$. That is, for every $(\phi(x), y)$ in $S'$, and
there exists a vector $u$ in $\fR^{d'}$, such that for every $(\phi(x), y)$ in $S'$, $\frac{y u \cdot \phi(x)}{\|u\|\| \phi(x)\|} \geq \gamma'$.
In addition, we would like our feature map $\phi$ to correspond to a kernel function $k(x,x') = \inner{\phi(x), \phi(x')}$ that can be efficiently evaluated.

\section{Construction of polynomials}
Our $\phi$'s in considerations are as follows: TODO: fix the constant in phi so that it has norm at most 1.
\begin{enumerate}
\item $\phi(x) = (x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq \deg, i_1, \ldots, i_d \geq 0} = (x_S)_{S \in N^d, \|S\|_1 \leq \deg}$. This corresponds to the complete symmetric kernel $K_{\deg}$.
\item $\phi(x) = ( \sqrt{ {\deg \choose i_1+\ldots+i_d} {i_1+\ldots+i_d \choose i_1 \ldots i_d}}  x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq \deg, i_1, \ldots, i_d \geq 0}$. This corresponds to the polynomial kernel
$K_{\deg}'(x,x') = (1 + \inner{x, x'})^{\deg}$.
\item $\phi(x) = ( \sqrt{ (\frac 1 2)^{i_1+\ldots+i_d} {i_1+\ldots+i_d \choose i_1 \ldots i_d}}  x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq \deg, i_1, \ldots, i_d \geq 0}$. This coresponds to the kernel $K(x,x') = \frac{1}{1 - \frac 1 2 \inner{x,x'}}$.
\end{enumerate}

For a polynomial $p = \sum_{S \in \fN^d} c_S x_S$, define $\|p\|$, the norm of $p$, as the $\ell_2$ norm of the vector that contains its coefficients on every monomial, that is $\|p\| = \| (c_S)_{S \in \fN^d} \| = \sqrt{\sum_{S \in \fN^d} c_S^2}$. This is also the RKHS norm of $p$ wrt the complete symmetric kernel $K_{\deg}$, for $\deg \geq \deg(p)$.

\subsection{Construction using rational functions}

\subsection{Construction using Chebyshev polynomials}
\paragraph{The construction.} Define
\[ p(x) = t + \frac 1 2 - \sum_{i=1}^t (T_{\ceil{\sqrt{\frac 2 \gamma}}}(1 - v_i \cdot x))^{\ceil{\log 2t}}, \]
where
$T_r(\cdot)$ is the $r$-th order Chebyshev polynomial. Denote by $C_i(x) = (T_{\ceil{\sqrt{\frac 2 \gamma}}}(1 - v_i \cdot x))^{\ceil{\log 2t}}$. $p$ can therefore be simplified to
\[ p(x) = t + \frac 1 2 - \sum_{i=1}^t C_i(x). \]

\begin{lemma}[~\cite{klivans2004learning}]
For all $(x,y)$ in $S$ satisfying Definition~\ref{def:int-margin}, we have that $y p(x) \geq \frac 1 2$.
\end{lemma}
\begin{proof}
We consider the two cases regarding the label of the example.
\begin{enumerate}
\item Suppose $y = 1$, then for all $i$ in $[t]$, $v_i \cdot x \geq \gamma$. This implies that for all
$i$ in $[t]$, $1 - v_i \cdot x \in [0, 1-\gamma]$. By item~\ref{item:bound-in} of Fact~\ref{fact:cheb}, we have that for all $i$, $|C_i(x)| \leq 1$.
Therefore, $p(x) \geq t + \frac 1 2 - t \geq \frac 1 2$.

\item Suppose $y = -1$. First observe that for all $i$, $T_{\ceil{\sqrt{\frac 2 \gamma}}}(1 - v_i \cdot x)) \geq -1$. This is because of items~\ref{item:bound-in} and~\ref{item:bound-out} of Fact~\ref{fact:cheb}.
This implies that $C_i(x) \geq -1$.
On the other hand, by Definition~\ref{def:int-margin}, there exists an $i_0$ in $[t]$, such that $v_{i_0} \cdot x \leq -\gamma$. By item~\ref{item:bound-out} of Fact~\ref{fact:cheb}, we have that
$C_{i_0}(x) \geq (1 + \gamma \cdot \frac 1 \gamma)^{\ceil{\log 2t}} \geq 2t$. Therefore,
\[ p(x) = t + \frac 1 2 - C_{i_0}(x) - \sum_{i \neq i_0} C_i(x) \leq t + \frac 1 2 - 2t + (t-1) \leq -\frac 1 2. \]
\end{enumerate}
\end{proof}

\begin{fact}
$T_r(\cdot)$ has the following properties:
\begin{enumerate}
  \item If $|z| \leq 1$, then $|T_r(z)| \leq 1$. \label{item:bound-in}
  \item If $z > 1$, then $T_r(z) \geq 1 + r^2 (z-1)$. \label{item:bound-out}
  \item $T_r(z)$ can be written as $\sum_{l=0}^r c_l z^l$, where coefficients $c_l$ are bounded by $2^r$ in absolute value.
\end{enumerate}
\label{fact:cheb}
\end{fact}

\paragraph{Norm bound.} We now bound $\| p \|$. First, for all $i$, as $v_i$ has norm at most 1,
\[ \| (1 - v_i \cdot x) \|^2 \leq 2. \]
Therefore, by Lemma~\ref{lem:norm-oper},
\[ \| c_l (1 - v_i \cdot x)^l \|^2 = c_l^2 \| (1 - v_i \cdot x)^l \|_2^2 \leq 2^{2r} l^l 2^l. \]
Hence,
\[ \| T_r(1 - v_i \cdot x) \|^2 = \| \sum_{l=0}^r c_l (1 - v_i \cdot x)^l \|^2 \leq (r+1) \cdot \sum_{l=0}^r 2^{2r} l^l 2^l \leq (r+1)^2 (8r)^r. \]
And by Lemma~\ref{lem:norm-oper},
\[ \| C_r(x) \|^2 \leq (\ceil{\log 2t})^{\ceil{\log 2t} r} \cdot ((r+1)^2 (8r)^r)^{\ceil{\log 2t}}. \]
Therefore,
\[ \| p(x) \|^2 \leq (t+1) \cdot ((t+\frac 1 2)^2 + t (\ceil{\log 2t})^{\ceil{\log 2t} r} \cdot ((r+1)^2 (8r)^r)^{\ceil{\log 2t}}) \leq 2 \cdot (32 r \ceil{\log 2t})^{2r \ceil{\log 2t}}  \]
where the last inequality uses the fact that $2^{2r \ceil{\log 2t}} \geq \max(2t^2, r^2)$.
\section{Auxiliary lemmas}
\begin{lemma}
   For $i=1,\ldots, \ell$ let $q^{(i)}(x)=\sum_S c^{(i)}_{S} x_S $ be a polynomial over $x_1, \ldots, x_d$. Then
   \begin{enumerate}
     \item if $q^{(1)}\ldots q^{(\ell)}$ has degree at most $\deg$, we have $\|\prod_{i=1}^{\ell} q^{(i)}\|^2 \leq \ell^{\deg} \prod_{i=1}^{\ell} \|q^{(i)}\|^2$. Specifically, for a polynomial $p$ of degree $s$, $\|p^\ell \|^2 \leq  \ell^{\ell s} (\|p\|^2)^\ell$.
     \item $\|\sum_{i=1}^{\ell} q^{(i)}\|^2 \leq \ell(\sum_{i=1}^{\ell} \|q^{(i)}\|^2)$.
   \end{enumerate}
   \label{lem:norm-oper}
\end{lemma}
